\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-22 notes/2loopNLSM.tex
final edits for ck at one-loop semiYM

\subsection{semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory.

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

As noted above, we can select out can manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of $(\varepsilon k)$ and $(kk)$. As we describe in detail in \cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
Y-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

\subsection{J-theory}
$J$-theory is obtained through a first order reformulation of NLSM in terms of the chiral current \cite{Cheung:2021zvb}.\footnote{For related formulations of NLSM that do not
treat color-kinematics duality see \cite{Freedman:1980us,
  Slavnov:1971mz}.} Two constraints are placed on the chiral current.  First, the field strength associated with $J_\mu$ should vanish,
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0\, ,
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge,
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0\, ,
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
Since both theories are in Lorenz gauge this means that \emph{J-theory and semi-abelian YM are identical}. Furthermore, this theory is one-to-one with the observables amplitudes of $YZ$-theory when only considering on-shell $Z$$\bar{Z}$ states. As such, the single vertex in this theory is the same as the pure vector vertex of $YZ$ theory in Lorenz gauge,
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)\, .
\end{equation}
By calculating an off-shell four-point correlation function it is possible to check that the $J$-theory, or semi-YM theory, Lagrangian automatically satisfies color-kinematics duality, as described in \cref{offShellCK}.
As might be expected, the kinematic algebra is the diffeomorphism algebra \cite{Cheung:2021zvb}. Note that the different external particles break manifest Bose symmetry.

>>>>

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory.

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$. Furthermore, a simple calculation shows that the off-shell four-point correlation function computed from this vertex satisfies the Jacobi identity of \cref{jacID}. This ensures that color-kinematics duality holds to all multiplicity at tree-level. Moreover, since the number of propagating degrees of freedom in $D$, rather than $D-2$ for gauge-invariant non-abelian vectors, taking a forward limit on the cubic graphs will yield valid color-dual representations at one-loop as well.

As noted above, we can select out the manifestly cubic
sector of semi-abelian Yang-Mills from the full theory of \cref{eq:ym-lag} by selecting only $(\varepsilon^+\varepsilon^+)$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level or by plugging in the on-shell states of $J$-theory \cite{Cheung:2020djz,Cheung:2021zvb} and
$YZ$-theory \cite{Cheung:2016prv}. Indeed, semi-YM is just the following building blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(A_1,...,\bar{A}_i,...,A_n)_{\text{tree}} = \sum_{i\neq j} (\varepsilon_i \varepsilon_j) \Delta^{(ij)}_{\text{tree}} \sim \sum_{i\neq j} (\varepsilon_i \varepsilon_j)  \sum_a \,[(\varepsilon k)^{n-2} (kk)^{3-n}]_a\,,
\end{equation}
and similarly so at one-loop
\begin{equation}
A(A_1,...,A_n)_{\text{1-loop}} = \Delta^{(\varnothing)}_{\text{1-loop}} \sim \sum_a\,[ (\varepsilon k)^{n} (kk)^{-n}]_a
\end{equation}
where $[\,\cdots]_a$ are terms with appropriate powers of $(\varepsilon k)$ and $(kk)$. As we describe in detail in \cref{sec:CKLagrangians}, to recover NLSM at one-loop we need to extract the $D$-dependent part
that corresponds to an internal $\bar{Y}Y$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed $YZ$-theory matches to NLSM for off-shell
$Y$-particles. We now provide explicit expressions for these one-loop amplitudes in the next section.

@@

\subsection{J-theory}
$J$-theory is obtained through a first order reformulation of NLSM in terms of the chiral current \cite{Cheung:2021zvb}.\footnote{For related formulations of NLSM that do not
treat color-kinematics duality see \cite{Freedman:1980us,
  Slavnov:1971mz}.} Two constraints are placed on the chiral current.  First, the field strength associated with $J_\mu$ should vanish,
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0\, ,
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge,
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0\, ,
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
Since both theories are in Lorenz gauge this means that \emph{J-theory and semi-abelian YM are identical}. Furthermore, this theory is one-to-one with the observables amplitudes of $YZ$-theory when only considering on-shell $Z$$\bar{Z}$ states. As such, the single vertex in this theory is the same as the pure vector vertex of $YZ$ theory in Lorenz gauge,
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)\, .
\end{equation}
By calculating an off-shell four-point correlation function it is possible to check that the Lagrangians of $J$-theory and semi-YM theory automatically satisfies color-kinematics duality, as described in \cref{offShellCK}.
As might be expected, the kinematic algebra is the diffeomorphism algebra \cite{Cheung:2021zvb}. Note that the different external particles break manifest Bose symmetry.
